QPLIB: A Library of Quadratic Programming Instances

### 2018-02-08 QPLIB-3.tex
update numbers in text of 3.3
- remove wrong parenthesis
- fixed some typos and a double-a

During the development of the library, a discussion ensued about
the expected goals that we wished to achieve. The following four goals
where finally identified:
%
\begin{enumerate}
 \item to represent as far as possible all the different categories of QP
       problems;
 %
 \item to gather ``challenging'' instances, i.e., ones which can not be easily
       solved by  state-of-the-art solvers;
 %
 \item to include, for each of the categories, a set of well-diversified
       instances; and
 %
 \item to obtain a set of instances which is neither too small, so as to
       preserve statistical relevance, nor too large so as to being
       computationally manageable.
\end{enumerate}
%
To achieve such goals, we performed the following two filters, applied in
a cascade.
%
\begin{itemize}
 \item \emph{First Instances Filter.}\\
       The first filter was designed to drastically reduce the number of
       instances by eliminating the ``easy'' ones. An empirical measure
       for the hardness of an instance is the CPU time needed by a
       complete solver (cf.~\S\ref{sec:algo}) to solve it to
       global optimality. Accordingly, for each of the gathered instance we
       ran the complete solvers in GAMS, which number depends on the category
       of the instance under consideration, cf.~Table \ref{t:solvers}.
       We then filtered according to a first
       measure of computational difficulty, i.e., we discarded all instances
       that are solved by at least 30\% of the complete solvers within a time
       limit of 30 seconds.
 %
 \item \emph{Second Instances Filter.}\\
       The goal of the second filter was to eliminate ``similar'' instances.
       We carefully analyzed the instances one by one, and we clustered them
       according to their features; for each cluster we kept only a few
       representatives, e.g. by eliminating all but a few of those with very similar size and coming from the same donor. Finally, in order to only
       keep computationally challenging instances we ran a a complete solver for QGQ with a time limit of 120 seconds; all the
       instances which have been solved to proven optimality within this time limit
       were discarded.
\end{itemize}
%
In Table~\ref{tab:filters} we summarize the two filter steps, which
allowed us to identify the final set of $319$ discrete instances and
$134$ continuous instances.

@@

\textcolor{red}{We now analyze the features of the instances selected to be part of the
 library. In Table \ref{tab:DDDDD}, we provide a global overview. The instances have been divided in \textit{continuous} vs \textit{discrete} and \textit{convex} vs \textit{nonconvex}, forming in this way, a  classification of 4 macro categories. As previously mentioned, an instance is classified \textit{discrete} if it contains at least a binary or integer variable, and \textit{continuous} otherwise. On the other hand, an instance is classified as \textit{nonconvex} if the objective function is nonconvex and/or at least one of the constraints is nonconvex, and \textit{convex} otherwise. }

@@

 The detailed characteristics of the instances are presented in Table
\ref{tab:DD} for \emph{discrete} instances
(\textit{*}\{\textit{B,M,I,G}\}\textit{*}) and in Table
\ref{tab:CC} for \emph{continuous} ones (\textit{*C*}).
For each category, the tables
report in column ``$\#$'' the corresponding number of instances. It can be seen
that the final set well respects the original distribution of the gathered
instances among the different categories. Indeed, the discrete categories
(\textit{LMQ}) or (\textit{QBL}) are well represented by $118$ and $59$
instances, respectively. Similarly, the continuous categories
(\textit{LCQ}) and (\textit{QCQ}) are well
represented by $50$ and $17$ instances, respectively. Moreover, the library
actually covers the large majority of all possible categories of instances.

>>>>

During the development of the library, a discussion ensued about
the expected goals that we wished to achieve. The following four goals
were finally identified:
%
\begin{enumerate}
 \item to represent as far as possible all the different categories of QP
       problems;
 %
 \item to gather ``challenging'' instances, i.e., ones which can not be easily
       solved by  state-of-the-art solvers;
 %
 \item to include, for each of the categories, a set of well-diversified
       instances; and
 %
 \item to obtain a set of instances which is neither too small, so as to
       preserve statistical relevance, nor too large so as to being
       computationally manageable.
\end{enumerate}
%
To achieve such goals, we performed the following two filters, applied in
a cascade.
%
\begin{itemize}
 \item \emph{First Instances Filter.}\\
       The first filter was designed to drastically reduce the number of
       instances by eliminating the ``easy'' ones. An empirical measure
       for the hardness of an instance is the CPU time needed by a
       complete solver (cf.~\S\ref{sec:algo}) to solve it to
       global optimality. Accordingly, for each of the gathered instance we
       ran the complete solvers in GAMS, which number depends on the category
       of the instance under consideration, cf.~Table \ref{t:solvers}.
       We then filtered according to a first
       measure of computational difficulty, i.e., we discarded all instances
       that are solved by at least 30\% of the complete solvers within a time
       limit of 30 seconds.
 %
 \item \emph{Second Instances Filter.}\\
       The goal of the second filter was to eliminate ``similar'' instances.
       We carefully analyzed the instances one by one, and we clustered them
       according to their features; for each cluster we kept only a few
       representatives, e.g. by eliminating all but a few of those with very similar size and coming from the same donor. Finally, in order to only
       keep computationally challenging instances we ran a complete solver for QGQ with a time limit of 120 seconds; all the
       instances which have been solved to proven optimality within this time limit
       were discarded.
\end{itemize}
%
In Table~\ref{tab:filters} we summarize the two filter steps, which
allowed us to identify the final set of $319$ discrete instances and
$134$ continuous instances.

@@

\textcolor{red}{We now analyze the features of the instances selected to be part of the
 library. In Table \ref{tab:DDDDD}, we provide a global overview. The instances have been divided in \textit{continuous} vs \textit{discrete} and \textit{convex} vs \textit{nonconvex}, forming in this way, a  classification of 4 macro categories. As previously mentioned, an instance is classified \textit{discrete} if it contains at least a binary or integer variable, and \textit{continuous} otherwise. On the other hand, an instance is classified as \textit{nonconvex} if the objective function is nonconvex (if minimization) or nonconcave (if maximization) and/or at least one of the constraints is nonconvex, and \textit{convex} otherwise. }

@@

 The detailed characteristics of the instances are presented in Table
\ref{tab:DD} for \emph{discrete} instances
(\textit{*}\{\textit{B,M,I,G}\}\textit{*}) and in Table
\ref{tab:CC} for \emph{continuous} ones (\textit{*C*}).
For each category, the tables
report in column ``$\#$'' the corresponding number of instances. It can be seen
that the final set well respects the original distribution of the gathered
instances among the different categories. Indeed, the discrete categories
\textit{LMQ} and \textit{QBL} are well represented by \textcolor{red}{134} and \textcolor{red}{91}
instances, respectively. Similarly, the continuous categories
\textit{LCQ} and \textit{QCQ} are well
represented by \textcolor{red}{52} and \textcolor{red}{30} instances, respectively. Moreover, the library
actually covers the large majority of all possible categories of instances.
